Pipit on the Post: Proving Pre- and Post-conditions of Reactive Systems

### 2024-01-15 2024pipit/section/60-Evaluation.tex
eval

We were able to verify the Kind2 implementation of the simplified trigger-fetch mechanism for trigger arrays containing up to 16 elements; above that, a 32-size array reported multiple runtime errors.
% Verification of a 4-size array took 11s of CPU time; an 8-size array took 13s; a 16-size array took 276s; and a 32-size array timed out after several hours and multiple runtime errors.
For reference, the M_TTCAN hardware implementation of TTCAN supports up to 64 triggers~\cite{bosch2019mttcan}.

We made a critical simplification in the Kind2 implementation, which was to modify the trigger-enabled set to be a single cycle index.
In the specification, the enabled set is implemented as a cycle-offset and repeat-factor.
Checking if a trigger is enabled in the current cycle requires nonlinear arithmetic, which is difficult for SMT solvers.
In our Pipit development, we can treat the definition of the cycle set abstractly.
However, in the Kind2 development, quantifiers cannot contain function calls, which means that we cannot hide the implementation of the enabled-set check by providing an abstract contract.

@@

\begin{figure}
  \center
\begin{tabular}{r|rr|rr|rr}
  & \multicolumn{4}{c|}{Kind2} & Pipit \\
  & \multicolumn{2}{c|}{simple enable-set} & \multicolumn{2}{c|}{full enable-set} & \\
  size & wall-clock & CPU & wall-clock & CPU & wall-clock & CPU \\
  \hline
  1 & 2s &  3s  & 6s & 12s & 6s & 6s \\
  2 & 3s &  3s  & 8s & 12s & 6s & 6s \\
  4 & 5s & 11s  & 12s & 24s & 6s & 6s \\
  8 & 8s & 13s  & 82s & 90s & 6s & 6s \\
  16 & 125s & 276s & \multicolumn{2}{c|}{(error)} & 6s & 6s \\
  32 & \multicolumn{2}{c|}{(error)} & \multicolumn{2}{c|}{(error)} & 6s & 6s \\
\end{tabular}
\caption{Verification time for trigger-fetch; simple enable-set uses a simplified version of the enable-set, while full enable-set uses bitwise arithmetic as in the TTCAN specification.}
\label{f:evaluation:kind2-runtime}
\end{figure}

>>>>

We were able to verify the Kind2 implementation of the simplified trigger-fetch mechanism for trigger arrays containing up to 16 elements; above that, a 32-size array reported multiple runtime errors and did not terminate after several hours.
% Verification of a 4-size array took 11s of CPU time; an 8-size array took 13s; a 16-size array took 276s; and a 32-size array timed out after several hours and multiple runtime errors.
For reference, the M_TTCAN hardware implementation of TTCAN supports up to 64 triggers~\cite{bosch2019mttcan}.

We made a critical simplification in the Kind2 implementation, which was to modify the trigger-enabled set to be a single cycle index.
In the specification, the enabled set is implemented as a cycle-offset and repeat-factor.
Checking if a trigger is enabled in the current cycle requires nonlinear arithmetic, which is difficult for SMT solvers.
In our Pipit development, we can treat the definition of the cycle set abstractly.
However, in the Kind2 development, quantifiers cannot contain function calls, which means that we cannot hide the implementation of the enabled-set check by providing an abstract contract.
This limitation also makes the specification quite unwieldy, as functions must be manually inlined.

@@

\begin{figure}
  \center
\begin{tabular}{r|rr|rr|rr}
  & \multicolumn{4}{c|}{Kind2} & Pipit \\
  & \multicolumn{2}{c|}{simple enable-set} & \multicolumn{2}{c|}{full enable-set} & \\
  size & wall-clock & CPU & wall-clock & CPU & wall-clock & CPU \\
  \hline
  1 & 2s &  3s  & 6s & 12s & 6s & 6s \\
  2 & 3s &  3s  & 8s & 12s & 6s & 6s \\
  4 & 5s & 11s  & 12s & 24s & 6s & 6s \\
  8 & 8s & 13s  & 82s & 90s & 6s & 6s \\
  16 & 125s & 276s & \multicolumn{2}{c|}{error} & 6s & 6s \\
  32 & \multicolumn{2}{c|}{error} & \multicolumn{2}{c|}{error} & 6s & 6s \\
\end{tabular}
\caption{Verification time for trigger-fetch; simple enable-set uses a simplified version of the enable-set, while full enable-set uses bitwise arithmetic as in the TTCAN specification.}
\label{f:evaluation:kind2-runtime}
\end{figure}
